# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

```{r readdata, echo=TRUE}
activity=read.csv("activity.csv")
```

## Distribution of total number of steps taken per day

```{r histogram, echo=TRUE}
hist(activity$steps, main="Distribution of total number of steps per day", xlab="Total number of steps per day")
```

## What is mean total number of steps taken per day?

```{r mean, echo=TRUE}
meansteps=mean(activity$steps, na.rm=TRUE)
mediansteps=median(activity$steps, na.rm=TRUE)
```
Mean number of steps: `r meansteps`

Median number of steps: `r mediansteps`

## What is the average daily activity pattern?

```{r plot, echo=TRUE}
means=aggregate(activity$steps~activity$interval, activity, mean)
plot(means[,1], means[,2], type="l", ylab="Average number of steps", xlab="Time of day")

```
## Imputing missing values

```{r missingvalues, echo=TRUE}

totmsngsteps=sum(is.na(activity$steps))
totmsngintv=sum(is.na(activity$interval))
totmsngdate=sum(is.na(activity$date))
```
The number of missing values in the step variable is `r totmsngsteps`.
The number of missing values in the interval variable is `r totmsngintv`.
The number of missing values in the date variable is `r totmsngdate`.

```{r imputevalues, echo=TRUE}

mergeMeans=merge(activity,means,by.x="interval",by.y="activity$interval")
mergeMeans$steps[is.na(mergeMeans$steps)] = mergeMeans[is.na(mergeMeans$steps),4]

hist(mergeMeans$steps, main="Distribution of total number of steps per day with missing values imputed", xlab="Total number of steps")
meansteps=mean(mergeMeans$steps)
mediansteps=median(mergeMeans$steps)

```

Mean number of steps with missing values imputed: `r meansteps`

Median number of steps with missing values imputed : `r mediansteps`

## Are there differences in activity patterns between weekdays and weekends?

```{r plotweekend, echo=TRUE}
daysweek=weekdays(as.Date(mergeMeans$date))
weekend=c("Saturday","Sunday")
weekenddata=daysweek %in% weekend
datadays=cbind(mergeMeans,daysweek,weekenddata)
weekendfactor=factor(datadays$weekenddata, labels=c("Weekday","Weekend"))
datadays2=cbind(datadays,weekendfactor)

means2=aggregate(datadays2$steps~datadays2$interval+datadays2$weekendfactor, datadays2, mean)

library(ggplot2)
names(means2) = c("interval","weekendfactor","steps")
qplot(interval,steps,data=means2,facets=.~weekendfactor, geom="line", xlab="Time of day", ylab="Average number of steps")
```

